%---------------------------------------------------------------------
\section{Supplementation}
\label{sec:supplements}

%---------------------------------------------------------------------
\subsection*{Overview}

No supplement covered in this section has been tested in a climbing-specific RCT with a functional primary outcome (redpoint grade, route completion, pulley injury incidence, or return-to-climbing time). All evidence tiers for climbing applications are therefore extrapolated from adjacent populations or from surrogate-biomarker protocols. Single-source claims carry an explicit footnote; unresolved claims are flagged \toverify.

%---------------------------------------------------------------------
\subsection{Study Summary Table}
\label{subsec:supptable}

{\footnotesize
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.2}
\begin{longtable}{>{\raggedright\arraybackslash}p{2.0cm}
                  >{\raggedright\arraybackslash}p{2.0cm}
                  >{\raggedright\arraybackslash}p{2.0cm}
                  >{\raggedright\arraybackslash}p{2.2cm}
                  >{\raggedright\arraybackslash}p{2.5cm}
                  >{\raggedright\arraybackslash}p{1.2cm}
                  >{\raggedright\arraybackslash}p{1.2cm}
                  >{\raggedright\arraybackslash}p{1.0cm}}
\caption{Peer-reviewed intervention and review studies cited in this section.
         P1NP = amino-terminal propeptide of type I procollagen (collagen synthesis surrogate).
         VISA-A = Victorian Institute of Sport Assessment -- Achilles tendinopathy score.
         OSS = Oxford Shoulder Score; SPADI = Shoulder Pain and Disability Index.
         FSR = fractional synthesis rate (connective tissue).
         RTD = rate of torque development.
         CMJ/SJ = countermovement/squat jump.
         $g$ = Hedges $g$; $d$ = Cohen's $d$; SMD = standardised mean difference.
         S = surrogate; F = functional.
         ES not reported = authors did not publish standardised effect size or 95\,\% CI.}\label{tab:suppstudy}\\
\toprule
\textbf{Authors} & \textbf{Design / $n$} & \textbf{Supplement / dose} & \textbf{Primary outcome} & \textbf{Effect size} & \textbf{S/F} & \textbf{Quality} & \textbf{Industry} \\
\midrule
\endfirsthead
\toprule
\textbf{Authors} & \textbf{Design / $n$} & \textbf{Supplement / dose} & \textbf{Primary outcome} & \textbf{Effect size} & \textbf{S/F} & \textbf{Quality} & \textbf{Industry} \\
\midrule
\endhead
\midrule
\multicolumn{8}{r}{\itshape Continued on next page}\\
\endfoot
\bottomrule
\endlastfoot

Shaw et al.\ \cite{SLBR+17}
&
RDB crossover; $n=8$ healthy males, $21{\pm}0.5$\,yr; 7-day washout.
&
5 or 15\,g gelatin $+$ 48\,mg vit.~C vs.\ placebo; 1\,h before 6\,min rope-skipping.
&
Serum P1NP AUC 4\,h; in-vitro engineered-ligament mechanics.
&
15\,g: ${\approx}2{\times}$ P1NP vs.\ placebo ($p{<}0.05$); ligament stiffness/collagen $\uparrow$. $d$/CI not reported.%
\footnote{Single-source: P1NP at 4\,h reported as $+153\,\%$ vs.\ baseline for 15\,g; tables not independently verified.}
&
\textbf{S} (P1NP; in-vitro ligament)
&
Low
&
Baar lab; gelatin donated; patent conflict.\\

\addlinespace

Lis \& Baar \cite{LB19}
&
RDB crossover; $n=10$ recreationally active males.
&
15\,g vit-C-enriched gelatin, HC, gummy, or placebo; 1\,h pre jump rope.
&
Serum P1NP at 4\,h; amino-acid profile.
&
${\approx}20\,\%$ P1NP rise (trend only, NS vs.\ placebo). ES/CI not reported.
&
\textbf{S} (P1NP; amino acids)
&
Low
&
Same lab as Shaw; no conflict.\\

\addlinespace

Praet et al.\ \cite{PPW+19}
&
DB placebo-controlled crossover pilot RCT; $n=20$ chronic Achilles tendinopathy; 6-month.
&
5\,g/day TENDOFORTE\textsuperscript{\textregistered} $+$ calf strengthening vs.\ placebo.
&
VISA-A at 3\,months; microvascularity.
&
VISA-A: $+12.6$ (95\,\%\,CI 9.7--15.5) vs.\ $+5.3$ (2.3--8.3); MCID ${\approx}6.5$. Microvascularity NS.
&
\textbf{F} (VISA-A); \textbf{S} microvasc.
&
Moderate (pilot)
&
\textbf{GELITA\,AG. Flag.}\\

\addlinespace

Khatri et al.\ \cite{KNC+21}
&
SR of 15 RCTs; healthy adults and clinical populations.
&
Collagen peptides 5--15\,g/day $+$ exercise, 3--12\,wk.
&
Collagen markers, joint pain, body composition.
&
Collagen $+$ exercise improved joint pain/function and synthesis markers; no additional MPS effect. Pooled ES not calculated.
&
\textbf{S} (markers); \textbf{F} (pain)
&
Moderate (SR; heterog.)
&
Some trials industry-funded.\\

\addlinespace

Buchalski et al.\ \cite{BJAL26}%
\footnote{Two independent source citations; DOI requires resolver verification before submission.}
&
SR; 8 RCTs; $n=257$ (246M:11F); 3--15\,wk.
&
Collagen peptides 5--30\,g/day $+$ ${\geq}50$\,mg vit.~C $+$ high-load training.
&
Tendon CSA; stiffness; Young's modulus.
&
3/4 studies CSA $\uparrow$ vs.\ placebo ($+11\,\%$ vs.\ $+4.7\,\%$); stiffness/modulus $\uparrow$ in 2/2 high-dose. GRADE A structural.
&
\textbf{S-structural} (CSA, stiffness; large tendons only)
&
Moderate (GRADE A)
&
Not declared.\\

\addlinespace

Nulty et al.\ \cite{NTD+24}%
\singlesource
&
RDB crossover; $n=8$ resistance-trained men, $49{\pm}8$\,yr.
&
0, 15, or 30\,g vit-C-enriched HC 1\,h pre-resistance exercise.
&
P1NP AUC 0--6\,h.
&
15\,g: $+40\,\%$ vs.\ 0\,g ($p{=}0.024$); 30\,g: $+76\,\%$ ($p{<}0.001$); $+26\,\%$ vs.\ 15\,g ($p{=}0.033$).
&
\textbf{S} (P1NP only)
&
Low (small $n$; surrogate)
&
Not declared.\\

\addlinespace

Sas-Nowosielski et al.\ \cite{SNWK21}
&
DB placebo-controlled; $n=15$ elite climbers (13M:2F); 4\,wk.
&
$\beta$-alanine 4.0\,g/day vs.\ placebo; divided doses.
&
Campus reaches; traverse moves-to-failure and time-to-failure.
&
Campus: $d{=}1.55$ ($p{=}0.002$); traverse $+51\,\%$ ($d{=}1.73$); TTF $+32$\,s ($d{=}1.44$). Hard traverse: NS.
&
\textbf{S-functional} (climbing-specific; no grade endpoint)
&
Moderate (climbing-specific; small $n$)
&
None.\\

\addlinespace

Sandford et al.\ \cite{SSWL18}
&
Multicentre DB RCT; $n=73$ rotator-cuff shoulder pain; 8\,wk.
&
MaxEPA (1.53\,g EPA + 1.04\,g DHA) vs.\ oleic acid placebo; 9 capsules/day.
&
OSS at 2\,months; SPADI at 3\,months.
&
OSS: $-0.1$ (95\,\%\,CI $-2.6$ to $2.5$; $p{=}0.95$, null). SPADI: $-8.3$ (95\,\%\,CI $-15.6$ to $-0.94$; $p{=}0.03$).
&
\textbf{F} (pain/disability; shoulder; null primary)
&
Good
&
None.\\

\addlinespace

Grgic et al.\ \cite{Grg22}
&
SR + meta-analysis; 16 studies; $n=353$ (34F).
&
Caffeine 1--7\,mg/kg vs.\ placebo; acute.
&
Isometric handgrip strength.
&
$d{=}0.17$ (95\,\%\,CI 0.10--0.23; $p{<}0.001$); 1--3\,mg/kg: $d{=}0.20$ (0.10--0.30).
&
\textbf{S} (grip; no climbing outcome)
&
Moderate
&
None.\\

\addlinespace

Ruiz-L\'opez et al.\ \cite{RLMAMR+26}
&
Triple-blind RCT crossover; $n=13$ male climbers.
&
3\,mg/kg caffeine vs.\ placebo; 60\,min pre-testing.
&
Pull-up 1RM/power; dead-hang time; grip strength; RFD.
&
All outcomes NS: pull-up $+3.3\,\%$ ($g{=}0.061$); grip $+2.4\,\%$ ($p{=}0.060$); dead-hang $-7.4\,\%$ ($p{=}0.162$).
&
\textbf{S} (dynamometric; no on-wall outcome)
&
Moderate (underpowered)
&
None.\\

\addlinespace

Berlanga et al.\ \cite{BLSML+23}
&
DB crossover; $n=10$ male sport climbers, $28.8{\pm}3.7$\,yr.
&
70\,mL beetroot juice (6.4\,mmol NO$_3^-$) vs.\ nitrate-depleted placebo; 150\,min pre-testing.
&
Half-crimp test; pull-ups to failure; grip strength; CMJ; SJ.
&
All outcomes NS: no significant effect on any measure. ES not reported.
&
\textbf{S} (climbing-specific dynamometry; no grade endpoint)
&
Moderate
&
None.\\

\addlinespace

Bischof et al.\ \cite{BMSK24}
&
SR $+$ meta-analysis; 19 RCTs; $n{=}768$; ${\geq}8$\,wk.
&
Collagen peptides (mostly 15\,g/day) $+$ exercise vs.\ placebo.
&
Tendon morphology; fat-free mass; strength.
&
Tendon CSA: SMD\,0.67 ($p{=}0.011$, 4 studies). FFM: SMD\,0.48. Strength: SMD\,0.19. GRADE \textbf{very low} certainty for tendon; moderate for FFM.
&
\textbf{S-structural} (large tendons; no climbing data)
&
GRADE very low (tendon); heterogeneity; high RoB
&
Not declared.\\

\addlinespace

Miyamoto et al.\ \cite{MIO+25}
&
RDB; $n{=}50$ sedentary young males; 16\,wk; no structured training.
&
10\,g/day collagen peptides vs.\ placebo; no exercise protocol.
&
Gastrocnemius stiffness; Achilles stiffness; rate of torque development.
&
Gastrocnemius stiffness $d{=}0.594$; Achilles stiffness $d{=}0.378$; RTD $d{=}0.525$ vs.\ placebo.
&
\textbf{S-structural} (no training; sedentary; large tendons)
&
Moderate; industry-funded
&
\textbf{Industry-funded. Flag.}\\

\addlinespace

Kirmse et al.\ \cite{KLV+24}
&
RDB; $n{=}21$ resistance-trained males; 14\,wk.
&
15\,g/day collagen peptides $+$ training vs.\ placebo.
&
Connective tissue protein synthesis (FSR; deuterated water tracer).
&
No between-group difference in connective tissue FSR. ES not reported.
&
\textbf{S} (connective tissue synthesis rate; surrogate)
&
Moderate
&
Not declared.\\

\end{longtable}
}% end footnotesize

\noindent\textit{Note: No study in Table~\ref{tab:suppstudy} used a climbing-specific primary functional outcome (redpoint grade, on-wall endurance, pulley injury incidence). All climbing-specific evidence tiers are extrapolations.}

%---------------------------------------------------------------------
\subsection{Hydrolyzed Collagen Peptides}
\label{subsec:collagen}

\subsubsection{Mechanism}

Type~I collagen constitutes 65--80\,\% of tendon and ligament dry weight, and is the principal structural protein of finger flexor tendons, annular pulleys (A2, A4), and skin dermis --- all high-stress tissues in rock climbing. Collagen triple-helix formation requires hydroxylation of proline and lysine residues by prolyl-4-hydroxylase and lysyl hydroxylase, both of which are ascorbic-acid-dependent dioxygenases; without adequate vitamin~C, hydroxylation is rate-limited and secreted collagen is thermally unstable and subject to intracellular degradation. \textbf{[Evidence-backed: established biochemistry; Myllyharju, 2003, \textit{Matrix Biol}]}

Hydrolyzed collagen peptides (enzymatically processed to ${\approx}2$--5\,kDa; lower molecular weight than thermal gelatin) are absorbed intact from the gut as Pro-Hyp and Hyp-Gly dipeptides, which accumulate in connective tissue and may act as substrates for \textit{de novo} collagen synthesis and/or as signaling molecules upregulating tenocyte collagen gene expression. \textbf{[Bioplausible: mechanism established in vitro (Shigemura et al., 2009, \textit{J Agric Food Chem}); direct in-vivo mechanistic confirmation in human tendon tissue is lacking]}

Intermittent mechanical loading of tendons during exercise increases local blood flow and tenocyte metabolic activity, creating a window during which circulating amino-acid precursors are preferentially incorporated. The Shaw (2017) protocol exploited this by timing collagen ingestion 1\,h pre-exercise to synchronise peak plasma amino-acid elevation with exercise-induced hyperaemia. \textbf{[Bioplausible: inferred from P1NP kinetics; direct tissue-incorporation measurements absent]}

\subsubsection{Shaw et al.\ (2017) --- Critical Appraisal}

Shaw et al.\ (\textit{Am J Clin Nutr} 105:136--143; \cite{SLBR+17}) used a randomised, double-blind, crossover design in n=8 healthy males (age $21{\pm}0.5$\,yr; no tendon pathology). Three conditions --- 5\,g gelatin\,$+$\,48\,mg vitamin~C, 15\,g gelatin\,$+$\,48\,mg vitamin~C, and placebo --- were each administered as a single dose 1\,h before 6\,min rope-skipping, repeated 3$\times$/day for 3 days, with a 7-day washout.

\textbf{Primary outcomes}: (1) Serum P1NP --- a systemic surrogate biomarker of type~I collagen synthesis rate, \textit{not} a direct measure of tendon or pulley collagen content; (2) mechanical properties (stiffness, collagen content) of in-vitro engineered ligament constructs incubated with participant serum --- a cell-culture functional proxy, not a measurement in the participants' own tendons.

\textbf{Results}: The 15\,g condition produced ${\approx}2{\times}$ greater P1NP AUC vs.\ placebo over 4\,h post-exercise ($p{<}0.05$). Engineered ligament stiffness and collagen content were significantly greater with post-15\,g serum vs.\ post-placebo serum. The 5\,g condition was intermediate. Formal Cohen's $d$ and 95\,\%\,CI were not reported; the 2$\times$ P1NP estimate is a within-person comparison from n=8 with low precision.

\textbf{Critical limitations}:
\begin{enumerate}[noitemsep]
  \item n=8, male-only: underpowered; female hormonal effects on collagen turnover (estrogen upregulates collagen synthesis) unaddressed.
  \item Surrogate endpoint: P1NP reflects systemic whole-body collagen synthesis and does not confirm preferential incorporation into finger pulleys or flexor tendons. No tendon biopsy data exist.
  \item Healthy tendons only: tendinopathic tissue has an altered tenocyte phenotype and remodeling dynamics; results cannot be directly extrapolated to injured A2/A4 pulleys.
  \item Acute single-dose design: no chronic dosing, MRI, ultrasound, or functional load-to-failure outcome was assessed.
  \item Gelatin $\ne$ hydrolyzed collagen peptides: gelatin (thermally denatured) has higher molecular weight fragments than enzymatically hydrolyzed collagen (${\approx}2$--5\,kDa). Lis \& Baar (2019) found comparable amino-acid responses when both were vitamin-C-enriched, suggesting functional equivalence under those conditions.
  \item No climbing-specific outcome: rope-skipping loads Achilles/patellar tendons, not finger flexor pulleys.
  \item Industry conflict: the Baar lab has received collagen-industry support; a co-author holds patents on engineered ligament technology. Not disqualifying, but warrants bias assessment.
\end{enumerate}

\begin{evidencebox}{Collagen Pre-Exercise: P1NP Surrogate Benefit (Shaw 2017) · \evidencebacked · ESS 4}
\textbf{Evidence-backed (surrogate biomarker only):} 15\,g gelatin $+$ 48\,mg vitamin~C taken 1\,h before intermittent loading acutely elevates systemic P1NP in healthy young males (${\approx}2{\times}$; $p{<}0.05$) and improves in-vitro engineered ligament mechanics. \textit{Endpoint: surrogate (P1NP biomarker; ex-vivo ligament stiffness). No direct human tendon structural outcome, no injury incidence, no return-to-climbing, no climbing performance endpoint. This does not constitute evidence that climbers heal pulleys faster, increase tendon stiffness, or reduce injury incidence.}
\end{evidencebox}

\subsubsection{Lis \& Baar (2019) Confirmation}

Lis \& Baar (\textit{IJSNEM} 29:526--531; \cite{LB19}): n=10 males, crossover; all active collagen conditions (vitamin-C-enriched gelatin, hydrolyzed collagen, gummy) increased circulating amino acids (glycine, proline, hydroxyproline) similarly; P1NP showed a ${\approx}20\,\%$ trend from baseline that did not reach statistical significance due to high variability. The study confirms the amino-acid-kinetic window (peak ${\approx}1$\,h post-ingestion) and suggests functional equivalence between gelatin and hydrolyzed collagen forms when vitamin~C is co-supplemented. It does not establish a statistically reliable P1NP effect at 15\,g hydrolyzed collagen alone; the non-significance limits claims of confirmed replication.

\subsubsection{Praet et al.\ (2019): Clinical Tendinopathy Data}

Praet et al.\ (\textit{Nutrients} 11:76; \cite{PPW+19}) tested 5\,g/day specific collagen peptides (TENDOFORTE\textsuperscript{\textregistered}) $+$ calf-strengthening vs.\ exercise-only in n=20 chronic Achilles tendinopathy patients (pilot, double-blind, crossover RCT, 6 months). VISA-A improved by $+12.6$ (95\,\%\,CI 9.7--15.5) in the collagen-first group vs.\ $+5.3$ (95\,\%\,CI 2.3--8.3) in the placebo-first group; the between-group difference exceeds the MCID of ${\approx}6.5$ points. Tendon microvascularity decreased similarly in both groups.

\textbf{Limitations}: Pilot scale (n=20); crossover design with potential carry-over effects; single commercial peptide brand (TENDOFORTE\textsuperscript{\textregistered}); \textbf{funded by GELITA AG} (industry-funded: flag); Achilles tendon loading differs substantially from finger flexor pulley loading in climbing; extrapolation to A2/A4 pulleys requires a tissue-type and loading-pattern leap.

\begin{bioplausbox}{Collagen Peptides + Loading for Connective Tissue Recovery in Climbers · \bioplausible · ESS 2}
\textbf{Bioplausible} for collagen peptides $+$ structured loading supporting connective tissue recovery in climbers, drawing on Praet et al.\ (2019). The closest available analog involves a different tendon, a lower dose, a different loading protocol, and an industry-funded trial. No direct evidence exists for climbing-specific pulley healing.
\end{bioplausbox}

\subsubsection{Recent Structural Evidence (2024--2025)}
\label{subsubsec:collrecent}

The collagen literature expanded significantly in 2024--2025, providing both supportive and null findings.

\textbf{Bischof et al.\ (2024)} conducted the most comprehensive meta-analysis to date (\cite{BMSK24}): 19 RCTs, $n{=}768$ adults, long-term collagen peptides $+$ training. Pooled results: tendon CSA SMD\,0.67 ($p{=}0.011$, 4 studies, $n{=}114$); fat-free mass SMD\,0.48; maximal strength SMD\,0.19. GRADE certainty: \textbf{very low for tendon morphology} (heterogeneity, risk of bias, indirectness, imprecision, possible publication bias); moderate for body composition. No effect on tendon stiffness or Young's modulus in pooled analysis. These data strengthen the structural case for large tendons but the very-low GRADE certainty and complete absence of finger-pulley or climbing data must be noted.

\textbf{Kirmse et al.\ (2024)} tested 15\,g/day collagen during 14 weeks of resistance training in $n{=}21$ trained males, measuring connective tissue fractional synthesis rates (FSR) via deuterated water tracer (\cite{KLV+24}). No between-group differences were found. This null result directly challenges the assumption that oral collagen peptides augment connective tissue synthesis rates in trained athletes.

\textbf{Miyamoto et al.\ (2025)} randomised $n{=}50$ sedentary young males to 10\,g/day collagen or placebo for 16 weeks without structured exercise (\cite{MIO+25}). The collagen group showed increased gastrocnemius stiffness ($d{=}0.594$), Achilles stiffness ($d{=}0.378$), and rate of torque development ($d{=}0.525$) vs.\ placebo. This is the first RCT showing stiffness adaptation with collagen alone in untrained individuals; the sedentary population, absence of loading protocol, and industry funding limit generalisability to trained climbers.

\textbf{Synthesis}: positive structural signals exist for large tendons (Bischof 2024; Miyamoto 2025), but GRADE certainty is very low, Kirmse 2024 shows no synthesis rate benefit in trained athletes, and no study addresses finger pulleys or climbing load patterns.

\subsubsection{Dose}

Shaw (2017) demonstrated a dose-response between 5\,g and 15\,g for P1NP elevation; 15\,g produced the strongest signal. Lis \& Baar (2019) used 15\,g for all active conditions. Khatri et al.\ (2021; systematic review, 15 RCTs) identified 5--15\,g/day as the studied range, with 15\,g consistently associated with the highest collagen-synthesis marker responses. Praet et al.\ (2019) showed clinical efficacy at 5\,g/day in a chronic tendinopathy context with structured loading, suggesting lower doses may be sufficient for chronic daily protocols. No RCT has tested doses $>$15\,g in young healthy athletes; evidence from Nulty et al.\ (2024; \cite{NTD+24}) in middle-aged men suggests 30\,g produces $+26\,\%$ greater P1NP AUC than 15\,g ($p{=}0.033$), but this has not been independently replicated.

\textbf{[Evidence-backed:]} 15\,g is the best-supported acute pre-exercise dose for collagen synthesis biomarker elevation. No evidence that doses $>$15\,g provide additional clinical benefit in young athletes.

\subsubsection{Timing Window}

Shaw (2017) and Lis \& Baar (2019) both timed collagen ingestion 1\,h before exercise to align the plasma amino-acid peak (${\approx}60$\,min post-ingestion for 15\,g) with exercise-induced tendon hyperaemia. Gastric emptying variability (affected by co-ingested food, hydration, and sympathetic tone) could shift the peak by ${\pm}15$--30\,min. Taking collagen 45--75\,min pre-session adequately approximates the Shaw timing window. Immediate pre-exercise ingestion ($<$15\,min) likely misaligns the amino-acid peak with early high-strain loading.

\textbf{[Bioplausible:]} 1\,h pre-exercise is mechanistically optimal; direct evidence that the 1\,h window is superior to, e.g., 45\,min or 90\,min, does not exist from RCT comparisons.

\subsubsection{The Vitamin~C Co-Ingestion Gap}
\label{subsubsec:vitc}

\begin{warningbox}{\textbf{Protocol Gap: Vitamin\textasciitilde{}C Absent from Collagen Shake}}
The user takes 15\,g hydrolyzed collagen peptides pre-session \textbf{without} vitamin~C co-ingestion. Shaw et al.\ (2017) and Lis \& Baar (2019) both used vitamin-C-enriched preparations (48\,mg ascorbic acid per dose). Without this co-factor, the evidence-backed biomarker benefit \textbf{does not cleanly transfer}.
\end{warningbox}

\textbf{Biochemistry}: Prolyl-4-hydroxylase requires ascorbic acid as an obligatory electron donor; in its absence the hydroxylation cycle stalls, and hydroxylation efficiency falls in proportion to ascorbate depletion. Under frank deficiency ($<$10\,mg/day), collagen triple-helix formation fails entirely (scurvy). At suboptimal non-scorbutic intakes (plasma ascorbate $<$23\,$\mu$mol/L), hydroxylation efficiency is proportionally reduced. \textbf{[Evidence-backed: established enzymology]}

\textbf{Shaw protocol context}: 48\,mg vitamin~C was co-ingested per dose. The US RDA for vitamin~C is 90\,mg/day for men and 75\,mg/day for women; 48\,mg is sub-RDA but was not intended as a pharmacological dose --- its purpose was to standardise local ascorbate availability during the post-ingestion window without confounding the collagen signal.

\textbf{Scenario analysis} for the user's protocol:

\begin{center}
\begin{tabular}{>{\raggedright\arraybackslash}p{3.2cm}
                >{\raggedright\arraybackslash}p{3.8cm}
                >{\raggedright\arraybackslash}p{5.5cm}}
\toprule
\textbf{Vitamin~C status} & \textbf{Plasma ascorbate} & \textbf{Expected collagen synthesis response} \\
\midrule
Adequate habitual diet (${\geq}5$ fruit/veg servings/day; ${\geq}100$\,mg/day dietary C) & ${\geq}50$\,$\mu$mol/L (replete) & Prolyl hydroxylase likely not substrate-limited; response may approximate Shaw protocol. Co-ingestion provides marginal additional ascorbate. \\
\addlinespace
Suboptimal intake (common in athletes in caloric restriction; $<$50\,mg/day dietary C) & ${<}23$\,$\mu$mol/L (sub-optimal) & Substrate (Pro, Gly, Hyp) elevated but hydroxylation rate-limited; collagen synthesis efficiency impaired; evidence-backed benefit does not apply. \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Critical gap}: No RCT has directly compared 15\,g collagen \textit{with} vs.\ \textit{without} co-ingested vitamin~C under controlled dietary ascorbate conditions. The claim that omitting vitamin~C definitively abolishes benefit is \textbf{[Bioplausible]}, not experimentally proven. However, the claim that the user's protocol replicates the Shaw design is \textbf{false}: the designs differ in a mechanistically important cofactor. The uncertainty is asymmetric --- co-ingesting 50--100\,mg vitamin~C has no downside (cost $<$\$0.01/day; no risk at this dose) and eliminates the bottleneck. \textbf{Absence of co-ingestion is a correctable gap with no counterargument.}

\textbf{[Bioplausible]:} Adding ${\approx}50$\,mg ascorbic acid to the collagen shake is the most evidence-aligned protocol modification for the user. Whether it is strictly necessary depends on habitual dietary vitamin~C status, which has not been measured. No factorial RCT has proven that vitamin~C co-ingestion is strictly required in athletes with replete dietary ascorbate.

\subsubsection{Claim: Collagen for Climbing Skin Recovery}

\textbf{Claim}: Oral hydrolyzed collagen peptides accelerate healing of flappers (partial-thickness skin avulsions), skin tears, and chronic fingertip skin thinning in climbers.

\textbf{Evidence}: Puškarić et al.\ (2024; \textit{Nutrients}, \cite{FLAGA+24})\footnote{\toverify: only two sources cite a skin-specific RCT; the DOI cited by other sources resolves to Fernández-Lázaro 2024 omega-3 review. The skin RCT is a single-source claim requiring independent DOI verification.} examined collagen (5\,g $+$ 80\,mg vitamin~C/day $\times$ 16\,wk) in aging women (n=87), showing improved dermis density and texture vs.\ placebo. This population (cosmetic aging) and outcome (dermal density) differ mechanistically from acute traumatic skin avulsion and callus reformation in climbers.

No peer-reviewed RCT has tested oral collagen peptides for climbing-specific skin injuries.

\begin{folklorebox}{Collagen Accelerates Healing of Climbing Flappers and Split Tips · \folkloreverified · ESS 1}
\textbf{Folklore:} ``Collagen accelerates healing of climbing flappers and split tips.'' Source: climbing community forums (r/climbharder), coach blogs (Hooper's Beta, Lattice Training social media), supplement marketing. No peer-reviewed climbing-specific or acute wound-healing evidence exists for this claim. Bioplausible substrate mechanism (collagen provision $\to$ dermal fibroblast activity) does not constitute evidence.
\end{folklorebox}

\subsubsection{Claim: Collagen for Pulley and Tendon Remodeling Under Active Climbing Load}

\textbf{Mechanism} \textbf{[Bioplausible]}: Tendons and annular pulleys are ${\approx}70\,\%$ type~I collagen by dry weight. During loading, tenocytes upregulate TGF-$\beta$1 and IGF-1 signaling, increasing collagen turnover. Net anabolic balance requires adequate substrate (proline, glycine, hydroxyproline) and co-factors (vitamin~C). Provision of 15\,g hydrolyzed collagen elevates circulating Gly ($+376$\,$\mu$mol/L), Pro ($+162$\,$\mu$mol/L), and Hyp ($+105$\,$\mu$mol/L) at ${\approx}1$\,h post-ingestion (Shaw, 2017); active climbing concurrently increases tendon perfusion. The anabolic-window model is mechanistically coherent. \textbf{[Bioplausible: mechanism plausible; no climbing RCT measures pulley or flexor tendon collagen content change attributable to oral supplementation]}

Buchalski et al.\ (2026; systematic review; \cite{BJAL26}) report that collagen 15--30\,g/day $+$ vitamin~C $+$ high-load training (${\geq}70\,\%$ 1RM resistance) increases tendon CSA ($+11\,\%$ vs.\ $+4.7\,\%$ placebo) and stiffness/modulus in 3/4 studies; GRADE A for structural outcomes. The larger Bischof et al.\ (2024) meta-analysis (\cite{BMSK24}) pools 19 RCTs and finds tendon CSA SMD\,0.67 but GRADE certainty of \textbf{very low} for tendon morphology. These data involve large tendons under resistance-training load, not small-diameter finger pulleys under climbing-specific isometric crimp load. Additionally, Kirmse et al.\ (2024; \cite{KLV+24}) found no connective tissue protein synthesis rate benefit in trained athletes, providing an important null counterweight.

\begin{bioplausbox}{Collagen for Pulley and Tendon Remodeling Under Climbing Load · \bioplausible · ESS 2}
\textbf{Bioplausible:} Collagen peptides $+$ vitamin~C $+$ structured loading may support type~I collagen anabolism in finger flexor tendons and A2/A4 pulleys. The Praet (2019) Achilles tendinopathy data and Bischof (2024)/Buchalski (2026) structural data are the closest available analogs, but GRADE certainty for tendon morphology is very low, and Kirmse (2024) provides a null counterweight for trained athletes. No climbing-specific RCT exists for pulley healing, injury prevention, or return-to-climbing outcomes.
\end{bioplausbox}

\subsubsection{User Protocol Assessment and Recommended Correction}

\textbf{Current protocol}: 15\,g hydrolyzed collagen peptides pre-climbing session, no vitamin~C, unspecified exact timing.

\begin{center}
\begin{tabular}{>{\raggedright\arraybackslash}p{3.5cm}
                >{\raggedright\arraybackslash}p{2.5cm}
                >{\raggedright\arraybackslash}p{3.2cm}
                >{\raggedright\arraybackslash}p{3.5cm}}
\toprule
\textbf{Parameter} & \textbf{User protocol} & \textbf{Evidence-backed protocol} & \textbf{Gap} \\
\midrule
Dose & 15\,g \checkmark & 15\,g (Shaw/Lis/Khatri) & None \\
\addlinespace
Vitamin~C & \textbf{None} & ${\geq}48$--50\,mg co-ingested & \textbf{Evidence-aligned gap} \\
\addlinespace
Timing & ``Before climbing'' (unspecified) & 45--75\,min pre-session & Clarify; risk if $<$15\,min \\
\addlinespace
Frequency & Session days only & Daily (Shaw: 3$\times$/day $\times$3\,d; Baar 2017 recommendation: daily) & Uncertain gap; rest-day dosing untested \\
\addlinespace
Loading stimulus & Rock climbing & Intermittent loading ${\geq}6$\,min at moderate--high intensity & Climbing provides adequate stimulus \checkmark \\
\addlinespace
Monitoring & Unknown & Periodic grip/pain VAS assessment & No monitoring \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Minimum corrective action}: Co-ingest 50--100\,mg ascorbic acid (e.g., 100\,mg tablet or a small glass of citrus juice) with the 15\,g collagen shake, taken 45--60\,min before session start. This aligns the protocol with Shaw et al.\ (2017) and eliminates the mechanistic bottleneck at prolyl hydroxylase.

%---------------------------------------------------------------------
\subsection{Creatine Monohydrate}
\label{subsec:creatine}

\textbf{Mechanism}: Creatine supplementation saturates intramuscular phosphocreatine (PCr) stores, accelerating ATP resynthesis during high-intensity efforts of $\leq$15\,s (PCr-dependent energy system). Secondary: creatine may augment muscle hypertrophy via cell volumisation and mTOR pathway activation. Relevant to climbing: maximal boulder moves, contact-strength efforts, campus board power, and repeated attempts with brief rest intervals. \textbf{[Evidence-backed: well-replicated mechanism]}

\textbf{Evidence}: Multiple meta-analyses document creatine's efficacy for muscular strength and power in resistance-trained individuals. Pashayee-Khamene et al.\ (2024; \cite{PKHA+24}), a GRADE-assessed meta-analysis, confirms fat-free mass and strength gains from creatine $+$ resistance training. No recent climbing-specific RCT has measured creatine effects on boulder problem performance, maximum hang time, or campus board output in trained climbers. \textbf{[Evidence-backed for general strength/power; Bioplausible for climbing]}

\textbf{Dose}: No loading phase required for saturation if 3--5\,g/day maintained for $\geq$28 days. If rapid saturation is desired: 20\,g/day $\div$ 4 doses $\times$ 5--7 days, then 3--5\,g/day maintenance. Creatine monohydrate is the only form with robust evidence; effervescent, ethyl ester, and buffered variants show no superiority.

\textbf{Climbing-specific caveat --- body-mass penalty}: The central practical concern for climbers is acute body-mass gain of ${\approx}$0.5--2\,kg from intracellular water retention (Pashayee-Khamene et al.\ 2024; \cite{PKHA+24}), which may impair strength-to-weight ratio in weight-sensitive disciplines (lead climbing, slabs, long traverses). The net effect for power-oriented bouldering disciplines may be positive, but this has not been tested in a controlled climbing trial. Forearm muscles have relatively lower creatine storage capacity compared with large muscle groups, so climbers may experience smaller mass gains than reported in bodybuilding literature. Community claims that ``creatine makes you too heavy to climb'' should be treated as a practical concern rather than a proven contraindication.

\begin{evidencebox}{Creatine: Strength and Power in Resistance-Trained Athletes · \evidencebacked · ESS 4}
\textbf{Evidence-backed} for general strength and power in resistance-trained athletes (multiple meta-analyses, including Pashayee-Khamene et al.\ 2024; \cite{PKHA+24}). \textbf{Bioplausible} for climbing performance improvement; the body-mass caveat is the central practical limitation and applies specifically to climbing. No climbing-specific performance evidence exists.
\end{evidencebox}

\begin{folklorebox}{Creatine Makes You Too Heavy to Climb · \folkloreverified · ESS 2}
\textbf{Folklore:} ``Creatine makes you too heavy to climb.'' Source: climbing community forums (r/climbharder), coaching blogs (Lattice Training), gym culture. The magnitude of mass gain (0.5--2\,kg water retention) and its effect on climbing-specific metrics remain empirically untested. This is a legitimate practical concern for weight-sensitive disciplines, not a proven contraindication.
\end{folklorebox}

%---------------------------------------------------------------------
\subsection{Omega-3 Fatty Acids (EPA/DHA)}
\label{subsec:omega3}

\textbf{Mechanism}: EPA and DHA compete with arachidonic acid for COX-2 and LOX enzymatic pathways, shifting eicosanoid and leukotriene profiles toward less pro-inflammatory mediators. Potential relevance to climbing: modulating tendon/ligament inflammation in overuse pathology (pulley tendinopathy, medial epicondylopathy), reducing training-induced DOMS, and supporting synovial joint health in finger DIP/PIP joints. \textbf{[Evidence-backed: biochemical mechanism; Calder, 2017, \textit{Eur J Clin Nutr}]}

\textbf{Best tendon-specific RCT evidence}: Sandford et al.\ (2018; \cite{SSWL18}), n=73 rotator-cuff-related shoulder pain, 2.57\,g/day EPA+DHA $\times$ 8 weeks $+$ exercise: no clinically meaningful OSS benefit at 2 months (mean difference $-0.1$, 95\,\%\,CI $-2.6$ to $2.5$; $p{=}0.95$); modest SPADI improvement at 3 months ($-8.3$, 95\,\%\,CI $-15.6$ to $-0.94$; $p{=}0.03$) of uncertain clinical significance. No direct benefit for structural tendon outcomes. \textbf{[Evidence-backed for systemic anti-inflammatory effect; Evidence-backed (null result) for tendinopathy structural/pain outcomes in this population]}

Wang et al.\ (2026; \cite{WLZ+26}) conducted a two-sample Mendelian randomisation study and found genetically higher omega-3 levels were associated with an $+18\,\%$ increase in Achilles tendinitis risk ($p{=}0.020$).\footnote{The reported association was nominal ($p{=}0.020$) and did not survive multiple-testing correction in the authors' own analysis; the paper states ``further research is needed to confirm this relationship.'' Mendelian randomisation cannot establish causality and is subject to genetic pleiotropy; interpret with caution.}

\textbf{Dose}: 2--3\,g/day combined EPA+DHA is the dose range used in most anti-inflammatory trials. No dose-response curve for tendon outcomes. Daily supplementation; timing is not acutely critical. Antiplatelet effect at $>$3\,g/day; anticoagulant interaction at high doses; fish oil peroxidation if stored improperly; algal sources available for vegetarians.

\begin{bioplausbox}{Omega-3 for Tendon Inflammation Modulation in Climbers · \bioplausible · ESS 3}
\textbf{Bioplausible} for modulation of tendon-related inflammation in climbers with overuse pathology. Current controlled evidence does not confirm clinically meaningful tendon-structural or injury-prevention benefit. The best available tendinopathy RCT (Sandford 2018, rotator cuff) shows no primary-outcome benefit. No climbing-specific RCT exists.
\end{bioplausbox}

\begin{toverifybox}{Omega-3 Strengthens Tendons or Prevents Pulley Injuries in Climbers · \toverify · ESS 0}
\textbf{\toverify:} Claims that omega-3 ``strengthens tendons'' or ``prevents pulley injuries'' in climbers. Only two sources identified (physiotherapy blogs; sports nutrition marketing); requires $\geq$3 independent climbing-community sources. Not supported by current controlled evidence (Sandford 2018 null result).
\end{toverifybox}

%---------------------------------------------------------------------
\subsection{Magnesium}
\label{subsec:magnesium}

\textbf{Mechanism}: Mg$^{2+}$ is cofactor for $>$300 enzymatic reactions, including Mg-ATP complex formation, muscle relaxation (antagonises Ca$^{2+}$ at the sarcoplasmic reticulum), and sleep architecture modulation (GABA-receptor agonism). Relevant to climbing: neuromuscular recovery between moves, sleep quality for training adaptation, and potentially cramp prevention. \textbf{[Evidence-backed: physiological role]}

\textbf{Evidence}: Magnesium supplementation corrects documented deficiency and may improve performance in deficient individuals; effects in magnesium-replete athletes are small or absent. The Garrison et al.\ (2020) Cochrane review (11 RCTs, $n{=}735$; \cite{GKK+20}) concluded that magnesium supplementation is \textit{unlikely} to provide clinically meaningful cramp prophylaxis to individuals with idiopathic skeletal muscle cramps (moderate-certainty evidence). For sleep, a specific magnesium bisglycinate (250\,mg elemental/day) trial reported Insomnia Severity Index improvement ($d{\approx}0.2$, CI reported) at 4 weeks in poor sleepers.\singlesource No climbing-specific RCT exists for magnesium effects on forearm pump, grip endurance, or training recovery. \textbf{[Bioplausible for sleep-mediated recovery benefit in athletes with suboptimal status; Evidence-backed (null) for cramp prevention in general populations]}

Climbers in caloric restriction, heavy sweating, and high training loads are at elevated risk of suboptimal (not frank deficiency) magnesium status. Serum magnesium is an insensitive biomarker; erythrocyte or ionised magnesium testing is more informative.

\textbf{Dose}: 200--400\,mg elemental magnesium/day (as glycinate, malate, or citrate; superior bioavailability vs.\ oxide). Evening dosing to support sleep benefit. Diarrhoea common at $>$400\,mg oxide.

\begin{bioplausbox}{Magnesium for Sleep and Neuromuscular Recovery in Suboptimal-Status Climbers · \bioplausible · ESS 2}
\textbf{Bioplausible} for sleep-mediated recovery benefit in climbers with suboptimal magnesium status. \textbf{Evidence-backed} for correcting documented deficiency. \textbf{Evidence-backed (null)} for cramp prevention in general populations.
\end{bioplausbox}

\begin{folklorebox}{Magnesium Prevents Forearm Pump or Finger Cramps in Climbers · \folkloreverified · ESS 0}
\textbf{Folklore:} Claims that magnesium ``prevents forearm pump'' or ``stops finger cramps on routes.'' Source: coaching forums; gym culture; climbing nutrition blogs (3+ independent sources). No targeted trial supports this in euvolaemic, well-fed climbers. The Cochrane evidence (\cite{GKK+20}) does not support routine magnesium as a cramp-prevention supplement in replete individuals.
\end{folklorebox}

%---------------------------------------------------------------------
\subsection{Beta-Alanine}
\label{subsec:betaalanine}

\textbf{Mechanism}: $\beta$-alanine is the rate-limiting precursor for muscle carnosine synthesis. Carnosine ($\beta$-alanyl-L-histidine; pKa ${\approx}$6.83) acts as an intramuscular H$^+$ buffer, attenuating acidosis during high-intensity glycolytic work. Forearm pump in rock climbing --- the accumulating burning sensation from sustained gripping on crux sections and power-endurance routes --- is mechanistically attributable to intramuscular H$^+$ accumulation from glycolysis; carnosine's buffering capacity targets precisely this mechanism. \textbf{[Bioplausible for climbing; mechanism Evidence-backed in general contexts]}

\textbf{Evidence}: Hobson et al.\ (2012, \textit{Amino Acids}; \cite{HSB+12}): meta-analysis of 15 manuscripts (57 exercise measures, 360 participants); $\beta$-alanine significantly increased exercise capacity for efforts of 60--240\,s (median ES\,=\,0.374, IQR 0.140--0.747). Trexler et al.\ (2015, ISSN position stand; \cite{TSRS+15}): mean performance improvement $+2.85\,\%$ (95\,\%\,CI 0.37--10.49). Saunders et al.\ (2017, \textit{Br J Sports Med}; \cite{SESA+16}): updated meta-analysis, k=40 studies, $n{=}1461$; overall ES\,=\,0.18 (95\,\%\,CI 0.08--0.28), more conservative than the Hobson median. The duration range of 1--4\,min corresponds to sustained pump sections and power-endurance sequences on sport climbing routes. \textbf{[Evidence-backed for 60--240\,s high-intensity exercise; ESS note: the Saunders 2017 update (ES\,=\,0.18) is more conservative than the frequently quoted Hobson 2012 median of 0.374]}

\textbf{Climbing-specific RCT}: Sas-Nowosielski et al.\ (2021; \cite{SNWK21}): $\beta$-alanine 4\,g/day $\times$ 4 weeks in n=15 elite climbers (double-blind, placebo-controlled). Campus-board reaches: $d{=}1.55$, $p{=}0.002$; easy traverse moves: $+51\,\%$, $d{=}1.73$; easy traverse time-to-failure: $+32$\,s, $d{=}1.44$. Hard traverse: no significant benefit. The large effect sizes should be interpreted cautiously given the small sample (n=15, with 1 excluded) and multiple endpoints without correction for multiplicity. \textbf{[Evidence-backed for specific climbing-relevant endurance tasks; small n; moderate quality]}

\textbf{Dose}: 3.2--6.4\,g/day in divided doses (0.8--1.6\,g per dose, every 3--4\,h) to minimise paresthesia. Carnosine loading requires 4--12 weeks at therapeutic doses for ${\approx}$60--80\,\% increase in muscle carnosine content.

\textbf{Caveat}: Paresthesia (harmless tingling, primarily face/scalp/hands) is near-universal at single doses $>$800\,mg; divided or slow-release formulations mitigate this. Performance benefit is most pronounced for efforts of 60--240\,s; submaximal long-duration routes with low anaerobic contribution may see less benefit. No benefit for maximal strength or power output.

\begin{evidencebox}{Beta-Alanine: Climbing Endurance Task Benefit (Sas-Nowosielski 2021) · \evidencebacked · ESS 7}
\textbf{Evidence-backed} for 60--240\,s high-intensity exercise (Hobson 2012 meta: median ES\,=\,0.374; \cite{HSB+12}; Saunders 2017 meta: ES\,=\,0.18, 95\,\%\,CI 0.08--0.28; \cite{SESA+16}) and for specific climbing-endurance tasks in elite climbers (campus board, easy traverse; Sas-Nowosielski 2021, $d = 1.44$--1.73, but $n{=}15$ only; \cite{SNWK21}). Dose: 3.2--6.4\,g/day divided $\times \geq$4 weeks. \textit{Endpoint: campus-board and traverse tasks are climbing-specific surrogate-functional hybrids; no redpoint grade, route completion rate, or injury incidence endpoint. Hard traverse showed no benefit. The observed effect is most compatible with short-to-moderate-duration high-intensity efforts, not all forms of climbing pump.}
\end{evidencebox}

\begin{folklorebox}{Beta-Alanine Eliminates Forearm Pump in Climbing · \folkloreverified · ESS 1}
\textbf{Folklore:} ``Beta-alanine eliminates pump'' in climbing. Source: climbing blogs; nutrition forums; coaching websites (3+ independent sources). Current data show improvement in specific ${\approx}$1--4\,min tasks, not elimination of pump; hard traverse performance showed no benefit in the only climbing RCT (Sas-Nowosielski 2021; \cite{SNWK21}).
\end{folklorebox}

%---------------------------------------------------------------------
\subsection{Caffeine}
\label{subsec:caffeine}

\textbf{Mechanism}: Non-selective adenosine receptor antagonist (A1, A2A). Reduces perceived exertion, enhances neuromuscular activation, increases pain tolerance, potentiates motor unit recruitment, and accelerates calcium release from the sarcoplasmic reticulum. Relevant to climbing: sustained isometric grip force, delayed onset of perceived forearm pump, acute power output on difficult moves. \textbf{[Evidence-backed]}

\textbf{Best general evidence}: Grgic et al.\ (2022, \textit{Clin Nutr ESPEN} 47:89--95; \cite{Grg22}): meta-analysis, k=16 studies, n=353; isometric handgrip strength: $d{=}0.17$ (95\,\%\,CI 0.10--0.23; $p{<}0.001$). Subgroup 1--3\,mg/kg: $d{=}0.20$ (95\,\%\,CI 0.10--0.30). Small but statistically significant ergogenic effect on a measurement directly analogous to climbing grip. \textbf{[Evidence-backed for isometric grip strength]}

Grgic et al.\ (2020, \textit{Br J Sports Med} 54:681--688; \cite{GGP+19}): umbrella review of 21 published meta-analyses (k=34 independent studies across meta-analyses); muscle strength ES = 0.20 (95\,\%\,CI 0.13--0.26); muscular endurance ES = 0.41 (95\,\%\,CI 0.24--0.58). \textbf{[Evidence-backed for general muscular strength and endurance]}

\textbf{Climbing-specific RCT}: Ruiz-L\'opez et al.\ (2026; \cite{RLMAMR+26}): triple-blind randomised crossover, n=13 male climbers, 3\,mg/kg caffeine vs.\ placebo, 60\,min pre-testing. Pull-up reps: $+3.3\,\%$ (95\,\%\,CI $-3.30$ to $4.37$; $g{=}0.061$); maximal grip: $+2.4\,\%$ ($p{=}0.060$; $g{=}0.120$); dead-hang: $-7.4\,\%$ ($p{=}0.162$). All outcomes non-significant; effect sizes trivial--small. Subjective energy and alertness increased ($p{<}0.05$). This is the first triple-blind climbing-specific caffeine RCT and is underpowered ($n{=}13$); it does not rule out a small real effect at higher doses or with different populations. \textbf{[Evidence-backed: 3\,mg/kg does not meaningfully improve proxy climbing performance at this sample size; null $\neq$ no effect, especially given low power]}

\textbf{Dose}: 3--6\,mg/kg body mass, 45--60\,min pre-session. Lower doses (3\,mg/kg) sufficient for endurance and perceived-exertion benefit with fewer side effects; higher doses (5--6\,mg/kg) may increase maximal strength/power but risk anxiety, tachycardia, and sleep disruption. Habitual consumers show attenuated response; 48--96\,h abstinence before target sessions partially restores the effect.

\textbf{Climbing-specific caveat}: Fear and anxiety are performance factors in climbing, particularly on serious routes and in competition. Caffeine increases sympathetic arousal and may exacerbate anxiety-driven performance decrements, tremor, and hesitation on high-commitment moves. Sleep disruption at doses $>$3\,mg/kg within 6\,h of sleep onset reduces recovery quality.

\begin{evidencebox}{Caffeine: Isometric Grip Strength and Endurance Benefit · \evidencebacked · ESS 4}
\textbf{Evidence-backed} for isometric handgrip strength ($d{=}0.17$, Grgic 2022; \cite{Grg22}) and general muscular endurance ($d{=}0.41$, Grgic 2020; \cite{GGP+19}). \textbf{Evidence-backed (null at 3\,mg/kg, $n{=}13$)} for climbing-specific pull-up/grip metrics (Ruiz-L\'opez 2026; \cite{RLMAMR+26}). Dose 3--6\,mg/kg 45--60\,min pre-session; individual tolerance and habituation are critical modifiers. \textit{Endpoint: dynamometric surrogate (grip strength, endurance metrics); no redpoint grade, route completion, or competition ranking outcome in any trial. The Ruiz-L\'opez 2026 climbing-specific RCT showed null results on all proxy climbing metrics at 3\,mg/kg, while increasing subjective alertness.}
\end{evidencebox}

%---------------------------------------------------------------------
\subsection{Dietary Nitrate (Beetroot Juice)}
\label{subsec:nitrate}

\textbf{Mechanism}: Dietary nitrate (NO$_3^-$) is reduced to nitrite and then nitric oxide (NO) via the enterosalivary circuit, promoting vasodilation, reducing the oxygen cost of submaximal exercise, and potentially attenuating forearm muscle deoxygenation during sustained gripping. Community interest in beetroot juice (BRJ) for climbing targets the forearm pump mechanism: improved muscle oxygenation could delay the onset of ischaemic pump. \textbf{[Bioplausible: mechanism Evidence-backed in endurance sports; application to climbing-specific forearm ischaemia is extrapolated]}

\textbf{Climbing-specific RCT}: Berlanga et al.\ (2023; \cite{BLSML+23}): double-blind crossover, $n{=}10$ male sport climbers (age $28.8{\pm}3.7$\,yr), 70\,mL concentrated BRJ (6.4\,mmol NO$_3^-$) vs.\ nitrate-depleted placebo, 150\,min pre-testing. Outcomes assessed: half-crimp test, pull-ups to failure, isometric handgrip, countermovement jump, squat jump. All outcomes non-significant; no effect of dietary nitrate on any climbing-relevant neuromuscular measure. \textbf{[Evidence-backed (null): dietary nitrate at this dose does not improve climbing-specific neuromuscular performance]}

\textbf{Limitations}: $n{=}10$; male-only; amateur climbers; neuromuscular (strength/power) outcomes only --- does not test sustained pump or time-to-failure on a pump-heavy route; single acute dose; 150-min pre-exercise timing may not be optimal.

\begin{bioplausbox}{Dietary Nitrate for Forearm Pump Attenuation in Climbing · \bioplausible · ESS 3}
\textbf{Bioplausible} for forearm oxygenation and pump attenuation via NO-mediated vasodilation. The only climbing-specific RCT (Berlanga 2023; \cite{BLSML+23}) showed null effects on neuromuscular outcomes ($n{=}10$); however, the study did not test the vasodilatory/pump mechanism directly. No RCT has measured forearm oxygen re-saturation kinetics or time-to-failure on a pump-specific climbing task after BRJ supplementation.
\end{bioplausbox}

\begin{folklorebox}{Beetroot Juice Reduces Forearm Pump in Climbing · \folkloreverified · ESS 1}
\textbf{Folklore:} ``Beetroot juice reduces pump and improves endurance on the wall.'' Source: climbing nutrition blogs; r/climbharder supplement threads; endurance sports crossover community (3+ independent sources). Current climbing-specific evidence does not support this claim.
\end{folklorebox}

%---------------------------------------------------------------------
\subsection{Sodium Bicarbonate}
\label{subsec:nabic}

\textbf{Mechanism}: NaHCO$_3$ supplementation increases extracellular bicarbonate buffering capacity, counteracting H$^+$ efflux from working muscle during high-intensity glycolytic efforts. This is mechanistically complementary to $\beta$-alanine (which targets intracellular carnosine buffering): the two together address both sides of the intracellular--extracellular H$^+$ gradient. NaHCO$_3$ is classified Group A (evidence-supported ergogenic) by the IOC 2018 consensus statement. Standard dose: 0.3\,g/kg body mass, 60--90\,min pre-exercise. \textbf{[Evidence-backed for high-intensity efforts 1--7\,min: multiple meta-analyses; no climbing-specific RCT]}

\textbf{Evidence}: Meta-analyses confirm performance improvements for efforts of 1--7\,min, the precise range of climbing pump sections, with effect sizes comparable to or larger than $\beta$-alanine in acute single-dose protocols. Primary practical limitation: GI side effects (nausea, cramping, diarrhoea) occur in a substantial proportion of users at 0.3\,g/kg; encapsulated or hydrogel delivery formats partially mitigate this. \textbf{[Evidence-backed for high-intensity 1--7\,min efforts in non-climbing populations; Bioplausible for climbing forearm pump]}

\begin{bioplausbox}{Sodium Bicarbonate for Forearm Pump Management in Climbing · \bioplausible · ESS 3}
\textbf{Bioplausible} for reducing H$^+$-driven forearm pump on pump-intensive sport climbing routes via extracellular buffering. Evidence-backed for 1--7\,min high-intensity efforts in non-climbing populations. No climbing-specific RCT exists. GI side effects are the primary practical barrier; testing outside competition is essential before race-day use.
\end{bioplausbox}

%---------------------------------------------------------------------
\subsection{What the Literature Cannot Currently Resolve}
\label{subsec:suppgaps}

The field lacks any climbing-specific RCTs with hard functional outcomes (redpoint performance, injury incidence, pulley imaging, return-to-climbing time). The current evidence base is strongest for general strength/power supplements (creatine, caffeine), moderate for collagen as an adjunct to tendon rehab in non-climbing tendinopathy (with very-low GRADE certainty for structural outcomes), and weakest for translating any of this into actual climbing performance or injury prevention.

\begin{enumerate}[noitemsep]

  \item \textbf{Collagen $+$ vitamin~C and pulley structure in climbers}: No study has measured A2/A4 pulley or finger flexor tendon CSA, stiffness, or load-to-failure as a function of oral collagen supplementation using high-resolution ultrasound, MRI, or biopsy in any climbing population. New large-tendon CSA data (Bischof 2024 meta; \cite{BMSK24}) strengthen the structural case but do not address finger pulleys or climbing load patterns. Kirmse (2024; \cite{KLV+24}) shows no synthesis rate benefit in trained athletes.

  \item \textbf{Minimum effective vitamin~C co-dose}: Shaw (2017) used 48\,mg; Lis (2019) used enriched preparations. No factorial RCT has compared collagen $\pm$ vitamin~C under controlled dietary ascorbate conditions, so the threshold above which co-ingestion is redundant for athletes with replete diets remains undefined.

  \item \textbf{Dose-response above 15\,g}: No RCT has compared 15\,g vs.\ 20\,g vs.\ 30\,g collagen with functional tendon outcomes in young healthy athletes. The 30\,g data (Nulty 2024; \cite{NTD+24}, middle-aged men) suggest an additional P1NP increment but have not been independently replicated.

  \item \textbf{Collagen and climbing skin repair}: No controlled trial has assessed time-to-heal for flappers, split tips, or chronic skin thinning in climbers receiving collagen supplementation vs.\ placebo under standardised wound-care conditions.

  \item \textbf{Beta-alanine and perceived pump in climbing}: The Sas-Nowosielski (2021; \cite{SNWK21}) RCT showed campus board and easy traverse improvements but no hard traverse benefit. Whether carnosine loading reduces perceived pump and improves on-route performance in trained sport climbers (where vascular occlusion, not only metabolic acidosis, contributes to pump) is not established.

  \item \textbf{Caffeine at doses $>$3\,mg/kg in climbing}: The Ruiz-L\'opez (2026; \cite{RLMAMR+26}) RCT used 3\,mg/kg only. Whether 5--6\,mg/kg produces clinically meaningful improvement in actual climbing performance (redpoint grade, route completion, competition ranking) without offsetting anxiety or tremor remains untested in climbing populations.

  \item \textbf{Omega-3 and finger tendon remodeling}: The available tendinopathy RCT (Sandford 2018; \cite{SSWL18}, rotator cuff) showed no primary-outcome benefit. No RCT has examined EPA/DHA effects specifically on finger flexor tendons or pulleys, or on A2 pulley injury recurrence rate.

  \item \textbf{Interaction effects}: No study has examined whether collagen $+$ vitamin~C interacts with creatine, beta-alanine, or caffeine in terms of tendon adaptation, injury risk, or climbing-specific performance in a factorial or combination design.

  \item \textbf{Dietary nitrate and forearm pump in climbing}: The Berlanga (2023; \cite{BLSML+23}) RCT tested neuromuscular outcomes only. No RCT has directly tested BRJ or dietary nitrate on forearm oxygen re-saturation kinetics, time-to-failure on a pump-specific climbing circuit, or subjective pump ratings in trained sport climbers.

  \item \textbf{Sodium bicarbonate for forearm pump in climbing}: No RCT has directly tested NaHCO$_3$ (0.3\,g/kg, 60--90\,min pre-session) on subjective pump ratings, time-to-failure on a simulated pump route, or blood lactate kinetics during a climbing-specific circuit in trained sport climbers. The combination of $\beta$-alanine $+$ NaHCO$_3$ (dual intracellular/extracellular buffering) has shown additive effects in some sporting contexts but has not been tested in climbing.

\end{enumerate}
